---
title: "Exploratory Data Analysis"
author: "STOR 320/520 Group PLACE_GROUP_NUMBER_HERE (Ex: STOR 320.406 Group 12)"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(geosphere)
library(tibble)
# Import Data Below
# read_csv()
```

# Creator: Brian Houston
```{r}
locations<-readr::read_csv("https://raw.githubusercontent.com/littlecapybaraa/stor320-Project/main/locations.csv")
divisions<-readr::read_csv("https://raw.githubusercontent.com/littlecapybaraa/stor320-Project/main/divisions.csv")
standings <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-02-04/standings.csv')
games <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-02-04/games.csv')
```

### Q1: What is the relationship between a team's last season performance and their next season performance?

##A: There is a linear realtion ship between last season performance and next season performance 

```{r}
standings2 = standings %>% select(team, year, wins) %>% arrange(team, year)
for (i in 2:nrow(standings2)) {
  if (standings2[i-1,1] == standings2[i,1]) {
    standings2[i,"prev_wins"] = standings2[i-1,3]
  }
}
standings3 = na.omit(standings2)
ggplot(standings3, aes(x = prev_wins, y = wins, color)) +
  geom_point() +
  geom_jitter()+geom_smooth(method="lm")
```

### Q2: How does the distance teams have to travel affect their performance?

##A: The distance do affect teams performance, when they travel longer distance, there is a higher chance the hometeam wins 

```{r}
locations = locations %>% mutate(Team = ifelse(Team == "Forty-Niners", "49ers", Team))
games2 = games %>% 
  left_join(locations, by = c("home_team_name" = "Team")) %>% 
  rename(home_latitude = latitude, home_longitude = longitude) %>%
  left_join(locations, by = c("away_team_name" = "Team")) %>% 
  rename(away_latitude = latitude, away_longitude = longitude) %>%
  mutate(did_home_win = as.integer(winner == home_team))
for (i in 1:nrow(games2)) {
  games2[i, "distance"] = distHaversine(c(games2[i,]$home_longitude, games2[i,]$home_latitude),
                                        c(games2[i,]$away_longitude, games2[i,]$away_latitude))
}
ggplot(games2, aes(x = distance, y = did_home_win)) +
  geom_point() +
  geom_jitter() +
  geom_smooth(method="glm", method.args = list(family=binomial))
```

# Interpreter: Yu Chen
```{r}
attendance <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-02-04/attendance.csv')
standings <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-02-04/standings.csv')
games <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-02-04/games.csv')
library(tidyverse)

```
### Q1: Does a team's defensive or offensive ranking matter more for their overall placement?

#higher offfensve and defensive ranking leads to more wins  

#higher defensive points leads to more wins 

#in general the more defensive points a team have, the lesser the wins they will get, because they spend more time on defending the offensive form their opponents.

#the higher the defensive ranking the lesser the points for defensive performance 

#comapre to the quality of thier opponents, the quality of the team itself matters more in total wins 

```{r}
ggplot(standings)+geom_point(aes(x=wins,y=points_for))+
  geom_smooth(aes(x=wins,y=points_for),method=lm)

ggplot(standings)+geom_point(aes(x=wins,y=points_against))+
  geom_smooth(aes(x=wins,y=points_against),method=lm)

ggplot(standings)+geom_point(aes(x=points_for,y=points_against))+
  geom_smooth(aes(x=points_for,y=points_against),method=lm)



```

#higher defensive points leads to more wins 
```{r}
standings %>%
  #group_by(wins) %>%
  
  ggplot(aes(x=points_for,y=points_against,fill=wins)) +
    geom_tile() +#geom_point(size = 2)+
  scale_fill_gradientn(colors=c("black","lightskyblue","white"))+
    
theme_dark()
    
#in general the more defensive points a team have, the lesser the wins they will get, because they spend more time on defending the offensive form their opponents. 
```
#higher offfensve and defensive ranking leads to more wins  
```{r}
standings %>%
  #group_by(wins) %>%
  
  ggplot(aes(x=offensive_ranking,y=defensive_ranking,fill=wins)) +
    geom_tile() +#geom_point(size = 2)+
  scale_fill_gradientn(colors=c("black","lightskyblue","white"))+
    
theme_dark()


```
#the higher the defensive ranking the lesser the points for defensive performance
```{r}
standings %>%
  #group_by(wins) %>%
  
  ggplot() + geom_point(aes(x=points_against,y=defensive_ranking)) #the higher the defensive ranking the lesser the points for defensive performance 
    
```


#comapre to the quality of thier opponents, the quality of the team itself matters more in total wins 
```{r}
standings %>%ggplot() + geom_point(aes(x=wins,y=strength_of_schedule))

standings %>%
  #group_by(wins) %>%
  
  ggplot(aes(x=simple_rating,y=strength_of_schedule,fill=wins)) +
    geom_tile() +#geom_point(size = 2)+
  scale_fill_gradientn(colors=c("black","lightskyblue","white"))+
    
theme_dark()
```
#Point differential (points_for - points_against) larger difference leads to more wins 
```{r}
ggplot(standings)+geom_point(aes(x=wins,y=points_differential))+
  geom_smooth(aes(x=wins,y=points_differential),method=lm)
```

### Q2: How well can a model predict whether a team made the playoffs using the team's regular season statistics and what statistics are most important in the model?

# Not finished 

# almost finished the training dataset
#selection model need to be determined
```{r}
head(attendance)
head(standings)
head(games)
```

```{r}
homegames=games%>%filter(year==2001)%>%count(home_team)%>%arrange(desc(n))

awaygames=games%>%filter(year==2001)%>%count(away_team)%>%arrange(desc(n))

```

```{r}
attendance2=attendance%>%select("team",team_name,year,total,home,away)
yearlyatt=attendance2[!duplicated(attendance2), ]
```

```{r}
comb12=yearlyatt%>%left_join(y=standings,by =c("year","team","team_name")) # combine df standings and df attendance, only yearly data 
comb12$teamname1=paste(comb12$team,sep = " ",comb12$team_name) # combine team_name and team to a new colunm 
comb12["winsathome"]=0 #Create a new colunm for home wins 
comb12
```

```{r}

comb12

```

## Q3 what teams win more when they are at home 

##on average,teams win more when they are at home, home wins contributed 60% percent of their total wins. 

```{r}
#games[games$year=="2000","home_team"][1,]==comb12[comb12$year=="2000","teamname1"][17,]
#as.character(games[games$year=="2000","home_team"][1,])
#as.character(comb12[comb12$year=="2000","teamname1"][17,])
#comb12[comb12$year=="2000","teamname1"][17,]
#comb12$teamname1=paste(comb12$team,sep = " ",comb12$team_name)
#comb12[comb12$year==2000&comb12$team_name==games[games$year==2000,"home_team"][1,],]
#comb12[comb12[comb12$year=="2000","teamname1"][,1]==as.character(games[games$year=="2000","home_team"][1,])&comb12$year=="2000",]
#p2=as.character(games[games$year=="2000","home_team"][1,])
#comb12[comb12$teamname1==as.character(games[games$year=="2000","home_team"][1,])&comb12$year=="2000",]
#comb12[comb12$year=="2000","teamname1"][,1]==as.character(games[games$year=="2000","home_team"][1,])
#comb12$teamname1==games[games$year=="2000","home_team"][1,]
#games[games$year=="2000","home_team"][1,]
#comb12$teamname1
#comb12[comb12$teamname1==as.character(games[games$year==2000,"home_team"][1,])&comb12$year==2000,"winsathome"]=1+comb12[comb12$teamname1==as.character(games[games$year==2000,"home_team"][1,])&comb12$year==2000,"winsathome"]
#comb12[comb12$teamname1==as.character(games[games$year==2000,"home_team"][1,])&comb12$year==2000,"winsathome"]
```

 
```{r}
#get the home wins for each team in each year and add them into the combined data 
for (j in 2000:2019){
for (i in 1:nrow(games[games$year==j,])){ if (games[games$year==j,"home_team"][i,]==games[games$year==j,"winner"][i,]){ comb12[comb12$teamname1==as.character(games[games$year==j,"home_team"][i,])&comb12$year==j,"winsathome"]=comb12[comb12$teamname1==as.character(games[games$year==j,"home_team"][i,])&comb12$year==j,"winsathome"]+1
}}}


comb12

sum(comb12["winsathome"])

```

```{r}
which(comb12$wins==0)  
comb12[c(265,550),] 
```

```{r}
#remove the NaN rows and get the winsashome ratio, there a two teams that didn't win once in some year
mean(comb12$winsathome[-c(265,550)]/comb12$wins[-c(265,550)]) #average of the winsathome ratio 

# on average,teams win more when they are at home, home wins contribute 60% percent of their total wins 
games%>%count(home_team==winner)

```
#split data into train and test 
```{r}
train_data=comb12[-(comb12$year=="2019"),]
train_data
test_data=
```

```{r}
library(bestglm)
library(gmodels)
library(MASS)
Full = glm(factor(sb_winner)~ .,family = binomial, data = train_data)
MSE = (summary(Full)$sigma)^2
none = glm(factor(sb_winner)~1, family = binomial,data=comb12)

e1.step.model.AIC <- step(none,scope = list(upper=Full), direction="both",test="Chisq", trace = F)
summary(e1.step.model.AIC)

plot(e1.step.model.AIC)
```


#still finding the largest homewin team 
```{r}
total_wins_at_home=tapply(comb12$winsathome, comb12$teamname1, FUN=sum)

total_wins_at_home=as.data.frame(total_wins_at_home)

total_wins_at_home

total_wins_at_home <- tibble::rownames_to_column(total_wins_at_home, "teams")
total_wins_at_home

total_wins_at_home%>%ggplot(aes(x=teams,y=total_wins_at_home))+
                              geom_bar(position="dodge",stat="identity",fill="#99CCFF") +
  theme(axis.text.x = element_text(angle = 270, vjust = 0.5, hjust=1))      #which team has the most home wins 



```
```{r}
total_wins=tapply(comb12$wins, comb12$teamname1, FUN=sum)

total_wins=as.data.frame(total_wins)

total_wins_HW=cbind(total_wins,total_wins_at_home)

total_wins_HW=cbind(total_wins_HW,total_wins_HW$total_wins_at_home/total_wins_HW$total_wins)

colnames(total_wins_HW) <-c("wins_total", "teamnames2","HWins","winRatio")


mean(total_wins_HW$winRatio)##HW/TTW rario sum of all the years 

total_wins_HW%>%ggplot(aes(x=teamnames2,y=winRatio,Fill=HWins))+
                              geom_bar(position="dodge",stat="identity",fill="#009999") +
  theme(axis.text.x = element_text(angle = 270, vjust = 0.5, hjust=1))+scale_fill_viridis_c() #which team has the highest hw ratio 
  


max_ind=which.max(total_wins_HW$winRatio)

total_wins_HW[max_ind,c("teamnames2","winRatio")]

total_wins_HW[total_wins_HW$teamnames2=="New England Patriots","winRatio"]
```
# the Minnesota Vikings has the highest home win ratio with 0.6606061

#the new england priots have the most win and home wins since 2000, with a ratio at 0.6582278  
```{r}

total_wins_HW
total_wins_HW%>%ggplot()+geom_bar(aes(x=teamnames2,y=wins_total),fill = "#FFCC00",position="dodge",stat="identity") + geom_bar(aes(x=teamnames2,y=HWins),position="dodge",stat="identity",fill="#CC3333")+
  theme(axis.text.x = element_text(angle = 270, vjust = 0.5, hjust=1))  


```

```{r}
total_wins_HW%>%ggplot(aes(x=winRatio))+
                              geom_boxplot() 




#win rario with dots/    win ratio within each years for each team 
```
#relation between ttlwins and HWW ratio
```{r}
total_wins_HW %>%
  #group_by(wins) %>%
  
  ggplot(aes(x=winRatio,y=wins_total)) + geom_point()+geom_smooth()
```


# Orator: Jason Hu

### Q1:  What is the relationship between attendance numbers and team standing 

##A: Higher attendance leads more wins 
```{r}
df <- dplyr::left_join(attendance,standings,by = c("year", "team_name", "team"))

df3 <- select(df, team,team_name,total,wins)
group_by(df3,team_name) %>%
  summarize(totalattendance=sum(total)/100000000,totalwins=sum(wins)) %>%
  ggplot(aes(totalattendance,totalwins)) +
  geom_point() +
  geom_smooth(method='lm')
```

### Q2: How does the amount teams score trend over time? 
#A: The total points gets higher over years 

#And average points tend to become higher at least several weeks in a year 
```{r}
# Check total points and if they went up over the years
df2 <- select(df, team,team_name,year,points_for)
df2[!duplicated(df2), ]
group_by(df2,year) %>%
  summarize(totalpoints=sum(points_for)) %>%
  ggplot(aes(year,totalpoints)) +
  geom_point()
# Note that the houston Texans joined the NFL in 2002
```
```{r}
# See if points scored change weekly
df4 <- select(games, year,week,pts_win,pts_loss)
games1 <- group_by(df4, year,week) %>%
  filter(week!="ConfChamp" & week!="Division" & week!="SuperBowl" & week!="WildCard")
# Check score change weekly for every year
games1$week = as.double(games1$week)
summarize(games1,weeklypoints=(sum(pts_win)+sum(pts_loss))) %>%
  filter(year==2007) %>%
  ggplot(aes(week,weeklypoints)) +
  geom_point()
  
# Check score change weekly for average of all years
games2 <- group_by(df4, week,year) %>%
  filter(week!="ConfChamp" & week!="Division" & week!="SuperBowl" & week!="WildCard")
games2$week = as.double(games2$week)
summarize(games2,weeklypoints=(sum(pts_win)+sum(pts_loss))) %>%
  summarize(weeklypoints=mean(weeklypoints),.groups="drop_last") %>%
  ggplot(aes(week,weeklypoints)) +
  geom_point()
```
###Q3 Does the home team have advantage over away teams?


##A yes 
```{r}
home <- select(games,year,home_team,away_team,winner)
home$homewin <- home$home_team==home$winner
home %>% group_by(year,homewin) %>% summarize(n=n()) %>%
ggplot(aes(x=year, y=n, fill=homewin)) +
  geom_bar(position="dodge",stat="identity")
```
same as the queation below

# Deliverer: Eleanna Janowski

### Q1: Does the home team have an advantage over away teams?

##A: Overall I think they have an a advantage over away team 
```{r}
games %>% mutate(did_home_win = winner == home_team) %>% 
  group_by(year) %>%
  summarize(n = n(), home_wins=sum(did_home_win),.groups = 'drop') %>%
  mutate(home_win_perc = home_wins/n) %>%
  ggplot(aes(x = year, y = home_win_perc)) +
  geom_line()
```

### Q2: How does division performance trend over time?

#A their performance tend to be closer in resent years 

```{r}
standings %>% left_join(divisions, by = c("team_name" = "Team")) %>%
  group_by(Division, year) %>%
  summarize(wins = sum(wins),.groups = 'drop') %>%
  ggplot(aes(x = year, y = wins, group = factor(Division), color = factor(Division))) +
  geom_line()
  
```
```{r}
standings %>% left_join(divisions, by = c("team_name" = "Team")) %>%
  group_by(Division, year) %>%
  summarize(wins = sum(wins),.groups = 'drop') %>%
  ggplot(aes(x = year, y = wins, group = factor(Division), color = factor(Division))) +
  geom_smooth(method="lm",se = FALSE)
```


# Follow-up Questions

### New Questions Based Off Initial Investigation

- Q1: what teams win more when they are at home 
- Q2: Are there trends in Team popularity based on performance through the years?
- Q3: How has home attendance for each team changed over time? 
- Q4: does team attendance turned down as they have a bad record? 

### Investigation of Follow-up Questions

GIVE WHAT 2 QUESTIONS YOU ATTEMPTED TO INVESTIGATE FURTHER IN COMPLETE SENTENCES (*Example:* Our group decided to investigate Q2 and Q4 in further detail.)

SHOW AT LEAST 2 TABLES OR FIGURES BELOW THAT EXPLORE ANSWERS FOR THE QUESTIONS YOU ARE INVESTIGATING FURTHER.


##Are there trends in Team popularity based on performance through the years?

#A:more home team win leads to more attendance 
```{r}
games %>% filter(!is.na(strtoi(week))) %>%
  mutate(weekn = as.double(week)) %>%
  arrange(home_team, year, weekn) %>%
  left_join(attendance, by = c("home_team_city" = "team", "weekn" = "week", "year" = "year")) %>%
  group_by(home_team) %>% 
  mutate(max_attendance = max(weekly_attendance)) %>%
  ungroup() %>%
  mutate(attendance_perc = weekly_attendance/max_attendance, did_home_win = as.integer(home_team == winner)) %>%
  ggplot(aes(x = attendance_perc, y = did_home_win)) +
  geom_point() +
  geom_jitter() +
  geom_smooth(method="glm", method.args = list(family=binomial))
```
voiline
# Does yearly attendance change based on winning or losing?

#A can't sey whether there is a realationship

```{r}
yeardf <- select(df,-week,-weekly_attendance)
yeardf <- yeardf[!duplicated(yeardf), ]
par(mfrow=c(2,2))
filter(yeardf,team_name=="Chargers") %>%
  ggplot(aes(x=year, y=total, fill=wins)) + 
  geom_bar(position="dodge",stat="identity")
filter(yeardf,team_name=="Chiefs") %>%
  ggplot(aes(x=year, y=total, fill=wins)) + 
  geom_bar(position="dodge",stat="identity")
filter(yeardf,team_name=="Cardinals") %>%
  ggplot(aes(x=year, y=total, fill=wins)) + 
  geom_bar(position="dodge",stat="identity")
```
plot it seprately
#Q What teams have the largest home attendance?

# there might be some errors in the code

#Redskins has the highest total attendence
```{r}
games$week <- as.double(games$week)
newdf <- left_join(games,attendance,by=c("year","week","home_team_name"="team_name")) %>% select(-team,-home,-total,-away) %>% rename(game_attendance=weekly_attendance)
newdf <- newdf[!is.na(newdf$game_attendance),]
```

```{r}

select(newdf,year,home_team_name,away_team,game_attendance) %>% group_by(year,home_team_name) %>% summarize(yearly_attendance=sum(game_attendance)) %>% 
  filter(year=="2018") %>%
  ggplot(aes(x=home_team_name, y=yearly_attendance,fill=home_team_name)) + 
  geom_bar(position="dodge",stat="identity") +
  theme(axis.text.x = element_text(angle = 270, vjust = 0.5, hjust=1))
#Total
select(newdf,year,home_team_name,away_team,game_attendance) %>% group_by(home_team_name) %>% summarize(total_attendance=sum(game_attendance)) %>% 
  ggplot(aes(x=home_team_name, y=total_attendance,fill=home_team_name)) + 
  geom_bar(position="dodge",stat="identity") +
  theme(axis.text.x = element_text(angle = 270, vjust = 0.5, hjust=1))
```
first 10- last 10 from highest to low 
# Summary

### Findings from first initial Questions:
  - Home teams win a noticeable amount more than away teams.
  - Teams have overall scored more points as years pass
  - Total attendance has a positive correlation with total wins for a team's overall lifespan

PARAGRAPH 1 

PARAGRAPH 2 SHOULD SUMMARIZE WHAT YOU LEARNED FROM INVESTIGATING THE FOLLOW-UP QUESTIONS. WHY ARE THESE FOLLOW-UP QUESTIONS INTERESTING FOR INVESTIGATION? DESCRIBE THE TABLES/FIGURES YOU USED TO EXPLORE ANSWERS TO THESE FOLLOW-UP QUESTIONS? WHAT DID YOU LEARN FROM THE TABLES/FIGURES REGARDING THE FOLLOW-UP QUESTIONS YOU PROPOSED?

### Findings from follow-up questions:
  - Small yearly trends where if a team did poorly total attendance lowers next year and vice versa











